### 2012-06-20 joint action and knowing others minds/submit philosophical studies  2012-01/second submit/interacting_mindreaders_source/interacting mindreaders.3.ps.corrections.tex
final version submitted to phil studies accepted

\def \papersize {a4paper}
\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

@@

\input{$HOME/Documents/submissions/preamble_steve_paper2}
%\author{}

@@

\begin{abstract}
Could interacting mindreaders be in a position to know things which they would be unable to know if they were manifestly passive observers?
This paper argues that they could.
Mindreading is sometimes reciprocal: 
the mindreader's target reciprocates by taking the mindreader as a target for mindreading.
The paper explains how such reciprocity can significantly narrow the range of possible interpretations of behaviour 
where mindreaders are, or appear to be, in a position to interact.
It follows that
revisions and extensions 
are needed to 
standard theories of mindreading
on which
all the evidence 
is in principle available to manifestly passive mindreaders.
The view also has consequences for 
understanding 
how abilities to interact combined with comparatively simple forms of mindreading
may explain the emergence,
in evolution or development,
of sophisticated forms of social cognition.

@@

Our question is part of a broader question,
What is the evidential basis for ascriptions of thought and action and how does the evidence support the ascriptions?
%\footnote{
%See e.g.\ \citet[pp.\ 126-7]{Davidson:1973jx}.
%}
%To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
%% nor directly about the mechanisms, processes or representations involved in mindreading
%, Davidson 
%sometimes formulates the question by asking what someone \emph{could} know that \emph{would} put them in a position to identify another's thoughts and actions.%
%\footnote{
%See, e.g., \citet[p.\ 126]{Davidson:1973jx}.
%While what follows draws  on Davidson's questions and insights,
%our aims here are more modest than his.
%For here we are concerned only with a  fraction of the problem of ascribing mental states and meanings,
%and we are not concerned with larger claims about mind, meaning and truth 
%(on which see 
%\citet{Davidson:1990du} and 
%\citet{lepore_donald_2005}).
%}
The most sustained attempts to answer this question,
Davidson's (\citeyear{Davidson:1984wh}; \citeyear{Davidson:1990du}), Lewis' (\citeyear{lewis:1974ri}) and Dennett's (\citeyear{Dennett:1987sf}),
do not exploit the possibility of interaction.
The evidence and principles they consider are available to manifestly passive observers.
So on their theories,
a purely passive mindreader observing from behind a one-way mirror
is on a par with
a mindreader who, actually or apparently, could interact with those she seeks to interpret.
The two are on a par in this sense:
in principle the same evidence could be available to each, and each can exploit the same principles in moving from evidence to ascriptions of thought and action.
Of course these theories are all compatible with the idea that interaction might be useful for mindreading in practice.  
But on these theories interaction makes no difference to what can in principle be known.
In this paper we aim to show
that mindreaders actually or apparently capable of interacting with their targets are at an advantage not only in practice but also in theory.
Their ascriptions could rely on evidence and principles which would be unavailable if they were entirely passive observers.

@@

A second qualification concerns the scope of mindreading.
In most discussions of mindreading, the focus is on  ascription of beliefs and other mental states to individuals.
But in what follows we shall focus on the ascription of goals to actions.
Some might claim that goal ascription is not mindreading, perhaps because 
identifying relations between actions and the outcomes to which they are directed
does not necessarily involve ascribing mental states.
Our view is consistent with this claim (and with its negation).
\label{goal_ascription_verify}
Evidence for ascriptions of belief and of other mental states often includes the occurrence (and non-occurrence) of goal-directed actions.
Identifying these actions typically requires goal ascription.
In some cases, then, 
evidence for the ascription of a goal
will indirectly support the ascription of a belief or other mental state---%
the evidence will support  a goal ascription which will in turn support the ascription of a belief (say) to the agent of the action.
Accordingly, even those who deny that goal ascription is mindreading should agree that sometimes evidence for goal ascription is indirectly evidence for ascription of belief and other mental states.
So whether or not goal ascription is deemed to be mindreading,
an individual's access to evidence for mindreading  depends in part on her access to evidence for goal ascription.
Our plan is to show that interacting mindreaders could make use, in ascribing goals, of evidence and principles which are not available to mere observers (who do not even appear to be in a position to interact with their targets).
In doing this we will be showing that
mindreaders who are (or appear) poised to interact with others 
could know things about their minds 
which they might not otherwise be in a position to know.

@@

While not decisive,
these considerations are perhaps sufficient to motivate 
exploring what evidence might support goal ascription.
According to 
what Csibra and Gergely call `the principle of rational action':
%
\begin{quote}
`an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality.'%
\footnote{
\citet[p.\ 255]{Csibra:1998cx}; cf.\ \citet{Csibra:2003jv}.
A related but different `principle of efficiency' has been formulated by \citet[p.\ 1061]{Southgate:2008el}:
`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.'
}
\end{quote}
%
Taking this idea as a rough starting point,
we propose that
%goal ascription could sometimes involve 
%identifying which potentially desirable outcomes 
%an observed action is the best available means to achieving.
%More carefully, our proposal is that 
these facts:
%
\begin{enumerate}
%
\item action $a$ is directed to some goal;
%
\item actions of $a$'s type are normally capable of being means of realising outcomes of $G$'s type in situations with the salient (to any concerned) features of this situation;
% 
\item no alternative type of action is both 
typically available to agents of this type 
and also 
such that actions of this type would be normally be significantly better\footnotemark \ means of realising outcome $G$ in situations with the salient features of this situation;
\footnotetext{
An action of type $a'$ is a better means of realising outcome $G$ in a given situation than an action of type $a$ if, for instance, actions of type $a'$ normally involve less effort than actions of type $a$ 
in situations with the salient features of this situation 
and everything else is equal; 
or if, for example, actions of type $a'$ are normally more likely to realise outcome $G$ than actions of type $a$
in situations with the salient features of this situation 
and everything else is equal.
}
%
\item the occurrence of outcome $G$ is typically desirable for agents of this type;
%
\end{enumerate}
%
and
%
\begin{enumerate}[resume]
\item there is no other outcome, $G'$, 
the occurrence of which would be at least comparably desirable for agents of this type 
and where (2) and (3) both hold of $G'$ and $a$
%
%where
%	the occurrence of $G'$ would be at least comparably desirable for agents of this type 
%%where
%%	$G$'s occurrence is not a means to $G'$'s nor conversely,
%and
%where  
%	(2) and (3) both hold of $G'$ and $a$
\end{enumerate}
%
may jointly constitute defeasible evidence for the conclusion that:
%
\begin{enumerate}[resume]
\item $G$ is a goal to which action $a$ is directed.
\end{enumerate}
%
We suggest that the above inference, from (1)-(5) to (6), is a route to knowledge of the goals of actions in this sense:
in some cases
it would be possible to know the premises without already knowing the conclusion;
and, 
in some of those cases,
knowing the premises could put one in a position to know the conclusion.%
\footnote{
Knowledge of the premises may not require explicit knowledge of 
	what is salient to others,
	what is desirable to them,
	or what makes actions better for them.
It is arguably sufficient that
the individual ascribing a goal 
is entitled to 
rely on being sufficiently similar to the target of her ascription 
with respect to 
these things.
%	what makes one action a better or worse means of realising an outcome than another action for her,  
%	what is desirable to her,
%	and what is salient to her.
}

@@

Our narrow aim in what follows is to show that some routes to knowledge of the goals of actions 
are available only to interacting mindreaders
and not to those who merely observe.
Having in this section introduced goal ascription as a topic,
the next step is to identify an obstacle to acquiring knowledge of the goals of actions,
one which we will eventually show can be avoided where mindreaders can (or appear to be able to) interact with their targets.

@@

That the problem of opaque means exists
is the first step in our argument
that 
theories of the evidential basis of mindreading based on pure observation 
are less powerful than theories taking into account the possibility of interaction.
They are less powerful in this sense: 
some routes to knowledge of the goals of actions are available only where a theory of the evidential basis of mindreading 
 takes the possibility of interaction into account.
In what follows we shall explain how abilities to interact with others provide a route to knowledge of the goals of other agents' actions  which avoids the problem of opaque means.
This is one way in which moving away from a purely observational model of mindreading yields a richer evidential base for ascriptions.

\section{Interactions involving distributive goals}
\label{sec:joint_action}
Our overall aim is to defend this claim:
there are routes to knowledge of the goals of others' actions
which are closed to
mindreaders who merely observe their targets
but open to 
mindreaders who are, or appear to be, capable of interacting with their targets.
As a preliminary to defending this claim,
we need to specify which types of interaction are relevant.
That is the aim of this section.

@@

To illustrate,
suppose that, while doing some gardening,
we find a turnip too big for any of us easily pull out of the ground alone.
We each individually intend that we all pull up the enormous turnip
and we act on this intention.%
\footnote{
For an argument that it is possible for each of us individually to intend that we do something, see \citet{Bratman:1999fr} and \citet[Chapter 2]{Bratman:2012fk}.
}
There is a single outcome,
freeing the turnip,
to which each of our actions is individually directed.
And it is possible for all of us to succeed in bringing about this outcome.
So this outcome is a distributive goal of our actions.

@@

For two agents' actions to have a distributive goal it is sufficient that their actions constitute a joint action 
(at least this is true on almost any account of joint action).%
\footnote{
\label{fn:joint_action_distributive_goal}
We know of two definitions of joint action on which it is not straightforward that all joint actions involve distributive goals.
One is provided 
 is provided by \citet[p.\ 366]{ludwig_collective_2007} and quoted in footnote \vref{fn:ludwig}; the other is due to 
\citet[p.\ 70]{Sebanz:2006yq}.
According to Sebanz and colleagues' `working definition', 
`joint action can be regarded as any form of social interaction whereby two or more individuals coordinate their actions in space and time to bring about a change in the environment'
(\citeyear[p.\ 70]{Sebanz:2006yq}).
On one way of reading this, the latter part is equivalent to `a change in the environment is such that: to bring it about two or more individuals coordinate their actions';
 this does suggest that the actions have a distributive goal.
On other standard accounts it is straightforward that joint action involves distributive goals.
Compare 
	\citet[pp.\ 329-31]{Bratman:1992mi},
	\citet[pp.\ 96-7]{Searle:1990em} and
	\citet[pp.\ 168-9]{gilbert:2009shared}.
}
However, the converse does not hold.
As the first variation of the turnip-pulling story indicates,
two or more agents' actions may have a distributive goal 
even though the agents do not know about each others' intentions or actions.
In fact our actions might have a distributive goal even though none of us is aware of this, or even of the others' existence. 
(We shall refrain from illustrating this point by contriving a truly gigantic turnip and a completely dark, very stormy night; readers can probably guess how this would go.)
One consequence of this is that 
two or more agents' actions may have a distributive goal
even though they are not engaged in joint action
(at least not on any standard account of joint action).%
\footnote{
On many accounts of joint action, each agent involved in a joint action must believe, expect or know something about the jointness of her action.
See 
	\citet[p.\ 103]{Bratman:1993je}, %common knowledge
	\citet[p.\ 40]{Butterfill:2011fk},
	\citet[p.\ 10]{Kutz:2000si}, %Agents acting on a shared intention have `a conception of themselves as contributors to a collective end.'
	\citet[p. 56]{miller_social_2001} % requires that each agent believes her actions are interdependent with the other agent's.
	and
	\citet[p.\ 361]{Roth:2004ki}. %: `each participant ... can answer the question of what he is doing or will be doing by saying for example ``We are walking together'' or ``We will/intend to walk together.''' 
Exceptions include  \citet{pacherie_framing_2011} and \citet{ludwig_collective_2007}.
\label{fn:ludwig}
According to Ludwig,
`The concept of a joint action as such is just that of an event of which there are multiple agents' (\citeyear[p.\ 366]{ludwig_collective_2007}).
Depending on what events are and what it is to be the agent of an event,
it may turn out that,
on Ludwig's definition,
our actions' having a distributive goal 
is sufficient for
us to be engaged in joint action.
}

@@

\section{Your-goal-is-my-goal}
\label{sec:your_goal_is_my_goal}
If a mindreader is able to interact with her targets,
if she is not limited to merely observing them,
how might this provide her with a route to knowledge of the goals of their actions?
The answer hinges on interactions involving distributive goals
(as defined \vpageref{df:distributive_goal}).

@@

Fortunately there is a way around this.  
For there are various cues which signal that one agent is prepared to engage in joint action with another,
and joint actions involve distributive goals.  
\label{twin_pram}
Seeing a stranger with the haggard look of a new parent struggling to get his twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.
(The noise of the street rules out talking.)   
In this way you signal that you are about to act jointly with the stranger, to lift the pram with him.   
Since the stranger is fully committed to getting his pram onto the bus,
he knows what the sole goal of his own actions will be.
But he also knows that you will engage in joint action with him,
which means that your actions and his actions will have a distributive goal.
This may enable him to infer the goal of your imminent actions: 
your goal is his goal, to get the pram onto the bus.

Our suggestion, then, is that the following inference characterises a route to knowledge of others’ goals:
%
\begin{enumerate}
\item You are willing to engage in some joint action\footnote{
We leave open the issue of how joint action is to be characterised subject only to the 
requirement that all joint actions must involve distributive goals (see further footnote \vref{fn:joint_action_distributive_goal}).
Attempts to characterise joint action in ways relevant to explaining development include
	\citet{Tollefsen:2005vh},
	\citet{Carpenter:2009wq},
	\citet{pacherie_framing_2011} and
	\citet{Butterfill:2011fk}.
The last of these 
not only shows how a primitive form of joint action can be characterised without ascriptions of higher-order mental states
but also shares our focus on relations between actions and goals rather than between agents and intentions.
}
or other with me
%(for example, because you have made eye contact with me while I was in the middle of attempting to do something).

@@

Is it ever possible to know the premises without first knowing the conclusion?  
Consider the first premise.  
Sometimes 
in the right contexts
an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she is about to engage in joint action with her.
Exploiting these indicators does not typically depend on knowing the particular contents of any beliefs, desires or goals.  
Expressions, gestures and movements can naturally indicate imminent jointness in much the way they can also naturally indicate emotions.\footnote{
Ideas along these lines are suggested by the discussion of {emergent coordination} in \citet{Knoblich:2010fk}.
}  
Of course these indicators provide no guarantee that others are genuinely willing to engage in joint action.  
But they are sufficiently reliable to ground knowledge in some cases.  
The existence of such indicators shows that knowing the first  premise of the above inference does not require already knowing which particular goals the other's actions will be directed to.

Not everything needs to rest on indicators, however.
It is sometimes possible to know that others are willing to engage in joint action with you even without relying on such indicators.  
Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are willing to act jointly.  
For example, this is often so for children struggling with a coat while surrounded by family or familiar adults.  
And in at least some subcultures people using public transport can reasonably take for granted that, within limits, those around them will act jointly with them when the need is clear.
Of course dispositions to engage in joint action may vary between cultural groups and situations.
This may be fatal for the peripatetic cosmopolitan,
but for others what matters is not whether the dispositions are universal but only that they are sufficiently widespread to be predictable.

Turning to the second point,
could knowing the premises of the your-goal-is-my-goal inference ever put one in a position to know the conclusion?  
Of course the inference is not deductive and will only work when certain background conditions are met.  
These background conditions include the other having largely true beliefs concerning which goal your actions are or will be directed to.
After all, the other agent may be willing to act jointly with you while being entirely mistaken about the goals to which your actions will be directed.
Where this happens, the premises of the inference might be true but the conclusion false.
(We return to this point in section \vref{sec:false_belief}.)  
But in some situations there is no requirement  to consider  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
Given the right conditions,
that others are willing to engage in joint action with you 
is sometimes sufficient reason to hold that
you will end up acting jointly with them
 even where the goal to which your actions will be directed is already fixed.

@@

@@

The problem of opaque means was this:
 failures to identify to which ends actions are means can impair goal ascription (see section \vref{sec:opaque_means}).
%As we saw, this problem is particularly likely to arise in cases involving novel tools and communicative gestures.
Showing how your-goal-is-my-goal makes  avoiding this problem possible is a way of demonstrating the potential value   of interaction for mindreading.

>>>>

%\def \papersize {a4paper}
\def \papersize {letterpaper}
\documentclass[12pt,\papersize]{extarticle}
% extarticle is like article but can handle 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, and 20pt text

@@

\input{preamble_steve_paper2}
%\author{}

@@

\begin{abstract}
\noindent
Could interacting mindreaders be in a position to know things which they would be unable to know if they were manifestly passive observers?  This paper argues that they could.  Mindreading is sometimes reciprocal: the mindreader's target reciprocates by taking the mindreader as a target for mindreading.  The paper explains how such reciprocity can significantly narrow the range of possible interpretations of behaviour where mindreaders are, or appear to be, in a position to interact.  A consequence is that revisions and extensions are needed to standard theories of the evidential basis of mindreading.  The view also has consequences for understanding how abilities to interact combined with comparatively simple forms of mindreading may explain the emergence, in evolution or development, of sophisticated forms of social cognition.

@@

Our question is part of a broader question,
What is the evidential basis for ascriptions of thought and action and how does the evidence support the ascriptions?
%\footnote{
%See e.g.\ \citet[pp.\ 126-7]{Davidson:1973jx}.
%}
%To stress that this question is not directly about about \emph{how} anyone actually ascribes thoughts or identifies actions% 
%% nor directly about the mechanisms, processes or representations involved in mindreading
%, Davidson 
%sometimes formulates the question by asking what someone \emph{could} know that \emph{would} put them in a position to identify another's thoughts and actions.%
%\footnote{
%See, e.g., \citet[p.\ 126]{Davidson:1973jx}.
%While what follows draws  on Davidson's questions and insights,
%our aims here are more modest than his.
%For here we are concerned only with a  fraction of the problem of ascribing mental states and meanings,
%and we are not concerned with larger claims about mind, meaning and truth 
%(on which see 
%\citet{Davidson:1990du} and 
%\citet{lepore_donald_2005}).
%}
The most sustained attempts to answer this question,
Davidson's (\citeyear{Davidson:1984wh}; \citeyear{Davidson:1990du}), Lewis' (\citeyear{lewis:1974ri}) and Dennett's (\citeyear{Dennett:1987sf}),
do not exploit the possibility of interaction.
The evidence and principles they consider are available to manifestly passive observers.
So on their theories,
a purely passive mindreader observing from behind a one-way mirror
is on a par with
a mindreader who, actually or apparently, could interact with those she seeks to interpret.
The two are on a par in this sense:
in principle the same evidence could be available to each, and each can exploit the routes to knowledge in moving from evidence to ascriptions of thought and action.
Of course these theories are all compatible with the idea that interaction might be useful for mindreading in practice.  
But on these theories interaction makes no difference to what can in principle be known.
In this paper we aim to show
that mindreaders actually or apparently capable of interacting with their targets are at an advantage not only in practice but also in theory.
Their ascriptions could exploit routes to knowledge which would be unavailable if they were entirely passive observers.

@@

A second qualification concerns the scope of mindreading.
In most discussions of mindreading, the focus is on  ascription of beliefs and other mental states to individuals.
But in what follows we shall focus on the ascription of goals to actions.
Some might claim that goal ascription is not mindreading, perhaps because 
identifying relations between actions and the outcomes to which they are directed
does not necessarily involve ascribing mental states.
Our view is consistent with this claim (and with its negation).
\label{goal_ascription_verify}
Evidence for ascriptions of belief and of other mental states often includes the occurrence (and non-occurrence) of goal-directed actions.
Identifying these actions typically requires goal ascription.
In some cases, then, 
evidence for the ascription of a goal
will indirectly support the ascription of a belief or other mental state---%
the evidence will support  a goal ascription which will in turn support the ascription of a belief (say) to the agent of the action.
Accordingly, even those who deny that goal ascription is mindreading should agree that sometimes evidence for goal ascription is indirectly evidence for ascription of belief and other mental states.
So whether or not goal ascription is deemed to be mindreading,
an individual's access to evidence for mindreading  depends in part on her access to evidence for goal ascription.
Our plan is to show that interacting mindreaders could exploit routes to knowledge of the goals of others' actions which are not available to mere observers.
In doing this we will be showing that
mindreaders poised to interact with others 
could know things about their minds 
which they might not otherwise be in a position to know.

@@

While not decisive,
these considerations are perhaps sufficient to motivate 
exploring what evidence might support goal ascription.
According to 
what Csibra and Gergely call `the principle of rational action',
%
\begin{quote}
`an action can be explained by a goal state if, and only if, it is seen as the most justifiable action towards that goal state that is available within the constraints of reality.'%
\footnote{
\citet[p.\ 255]{Csibra:1998cx}; cf.\ \citet{Csibra:2003jv}.
A related but different `principle of efficiency' has been formulated by \citet[p.\ 1061]{Southgate:2008el}:
`goal attribution requires that agents expend the least possible amount of energy within their motor constraints to achieve a certain end.'
}
\end{quote}
%
Taking this idea as a rough starting point,
we propose that
%goal ascription could sometimes involve 
%identifying which potentially desirable outcomes 
%an observed action is the best available means to achieving.
%More carefully, our proposal is that 
these facts:
%
\begin{enumerate}
%
\item action $a$ is directed to some goal;
%
\item actions of $a$'s type are normally capable of being means of realising outcomes of $G$'s type in situations with the salient (to any concerned) features of this situation;
% 
\item no alternative type of action is both 
typically available to agents of this type 
and also 
such that actions of this type would be normally be significantly better\footnotemark \ means of realising outcome $G$ in situations with the salient features of this situation;
\footnotetext{
An action of type $a'$ is a better means of realising outcome $G$ in a given situation than an action of type $a$ if, for instance, actions of type $a'$ normally involve less effort than actions of type $a$ 
in situations with the salient features of this situation 
and everything else is equal; 
or if, for example, actions of type $a'$ are normally more likely to realise outcome $G$ than actions of type $a$
in situations with the salient features of this situation 
and everything else is equal.
}
%
\item the occurrence of outcome $G$ is typically desirable for agents of this type;
%
\end{enumerate}
%
and
%
\begin{enumerate}[resume]
\item there is no other outcome, $G'$, 
the occurrence of which would be at least comparably desirable for agents of this type 
and where (2) and (3) both hold of $G'$ and $a$
%
%where
%	the occurrence of $G'$ would be at least comparably desirable for agents of this type 
%%where
%%	$G$'s occurrence is not a means to $G'$'s nor conversely,
%and
%where  
%	(2) and (3) both hold of $G'$ and $a$
\end{enumerate}
%
may jointly constitute defeasible evidence for the conclusion that:
%
\begin{enumerate}[resume]
\item $G$ is a goal to which action $a$ is directed.
\end{enumerate}
%
We suggest that the above inference, from (1)-(5) to (6), is a route to knowledge of the goals of actions in this sense:
in some cases
it would be possible to know the premises without already knowing the conclusion;
and, 
in some of those cases,
knowing the premises could put one in a position to know the conclusion.%
\footnote{
Knowledge of the conclusion may not require explicit knowledge of 
	what is salient to others,
	what is desirable to them,
	or what makes actions better for them.
It is arguably sufficient that
the individual ascribing a goal 
is entitled to 
rely on being sufficiently similar to the target of her ascription 
with respect to 
these things.
%	what makes one action a better or worse means of realising an outcome than another action for her,  
%	what is desirable to her,
%	and what is salient to her.
}

@@

Our narrow aim in what follows is to show that some routes to knowledge of the goals of actions 
are available only to interacting mindreaders
and not to those who merely observe.
Having in this section introduced goal ascription as a topic,
the next step is to identify an obstacle to acquiring knowledge of the goals of actions,
one which we will eventually show can be avoided where mindreaders can (or appear able to) interact with their targets.

@@

That the problem of opaque means exists
is the first step in our argument
that 
theories of the evidential basis of mindreading based on pure observation 
are less powerful than theories taking into account the possibility of interaction.
They are less powerful in this sense: 
some routes to knowledge of the goals of actions are available only where a theory of the evidential basis of mindreading 
 takes the possibility of interaction into account.
In what follows we shall explain how being able to interact with another sometimes makes available a route to knowledge of the goals of her actions  which avoids the problem of opaque means.
This will support our claim that an interacting mindreader might be in a position to know things which she would be unable to know if she were only observing.

\section{Interactions involving distributive goals}
\label{sec:joint_action}
We aim eventually is to defend this claim:
there are routes to knowledge of the goals of others' actions
which are closed to
mindreaders who merely observe their targets
but open to 
mindreaders who are, or appear to be, capable of interacting with their targets.
As a preliminary to defending this claim,
we need to specify which types of interaction are relevant.
That is the aim of this section.

@@

To illustrate,
suppose that, while doing some gardening,
we find a turnip too big for any of us to easily pull out of the ground alone.
We each individually intend that we all pull up the enormous turnip
and we act on this intention.%
\footnote{
For an argument that it is possible for each of us individually to intend that we do something, see \citet{Bratman:1999fr} and \citet{Bratman:2012fk}.
}
There is a single outcome,
freeing the turnip,
to which each of our actions is individually directed.
And it is possible for all of us to succeed in bringing about this outcome.
So this outcome is a distributive goal of our actions.

@@

For two agents' actions to have a distributive goal it is sufficient that their actions constitute a joint action 
(at least this is true on almost any account of joint action).%
\footnote{
\label{fn:joint_action_distributive_goal}
We know of two definitions of joint action on which it is not straightforward that all joint actions involve distributive goals.
One is provided 
 is provided by \citet[p.\ 366]{ludwig_collective_2007} and quoted in footnote \vref{fn:ludwig}; the other is due to 
\citet[p.\ 70]{Sebanz:2006yq}.
According to Sebanz and colleagues' `working definition', 
`joint action can be regarded as any form of social interaction whereby two or more individuals coordinate their actions in space and time to bring about a change in the environment'
(\citeyear[p.\ 70]{Sebanz:2006yq}).
On one way of reading this, the latter part is equivalent to `a change in the environment is such that: to bring it about two or more individuals coordinate their actions';
 this does suggest that the actions have a distributive goal.
On other standard accounts it is straightforward that joint action involves distributive goals.
Compare 
	\citet[pp.\ 329-31]{Bratman:1992mi},
	\citet[pp.\ 96-7]{Searle:1990em} and
	\citet[pp.\ 168-9]{gilbert:2009shared}.
}
However, the converse does not hold.
As the first variation of the turnip-pulling story indicates,
two or more agents' actions may have a distributive goal 
even though the agents do not know about each others' intentions or actions.
In fact our actions might have a distributive goal even though none of us is aware of this, or even of the others' existence. 
(We resist the temptation to contrive a truly gigantic turnip and a pitch dark, very stormy night; readers can probably guess how this would go.)
One consequence of this is that 
two or more agents' actions may have a distributive goal
even though they are not engaged in joint action
(at least not on any standard account of joint action).%
\footnote{
On many accounts of joint action, each agent involved in a joint action must believe, expect or know something about the jointness of her action.
See 
	\citet[p.\ 103]{Bratman:1993je}, %common knowledge
	\citet[p.\ 40]{Butterfill:2011fk},
	\citet[p.\ 10]{Kutz:2000si}, %Agents acting on a shared intention have `a conception of themselves as contributors to a collective end.'
	\citet[p. 56]{miller_social_2001} % requires that each agent believes her actions are interdependent with the other agent's.
	and
	\citet[p.\ 361]{Roth:2004ki}. %: `each participant ... can answer the question of what he is doing or will be doing by saying for example ``We are walking together'' or ``We will/intend to walk together.''' 
Exceptions include  \citet{pacherie_framing_2011} and \citet{ludwig_collective_2007}.
\label{fn:ludwig}
According to Ludwig,
`The concept of a joint action as such is just that of an event of which there are multiple agents' (\citeyear[p.\ 366]{ludwig_collective_2007}).
Depending on what events are and what it is to be the agent of an event,
it may turn out that,
on Ludwig's definition,
our actions' having a distributive goal 
is sufficient for
us to be engaged in joint action.
}

@@

\section{Your-goal-is-my-goal}
\label{sec:your_goal_is_my_goal}
If a mindreader is able to interact with her targets,
if she is not limited to merely observing them,
how might this enable her to exploit a route to knowledge of the goals of their actions?
The answer hinges on interactions involving distributive goals
(as defined \vpageref{df:distributive_goal}).

@@

Fortunately there is a way around this.  
For there are various cues which signal that one agent is prepared to engage in some joint action or other with another,
and joint actions involve distributive goals.  
\label{twin_pram}
Seeing you struggling to get your twin pram onto a bus and noticing you have the haggard look of a new parent, a passing stranger  grabs the front wheels and makes eye contact with you, raising her eyebrows and smiling.
(The noise of the street rules out talking.)   
In this way she signals that she is about to act jointly with you.   
Since you are fully committed to getting your pram onto the bus,
you know what the sole goal of your own actions will be.
But you also know that the stranger will engage in joint action with you,
which means that, taken together, her actions and your actions will have a distributive goal.
This may enable you to infer the goal of the stranger's imminent actions: 
her goal is your goal, to get the pram onto the bus.

Our suggestion, then, is that the following inference characterises a route to knowledge of others’ goals:
%
\begin{enumerate}
\label{your_goal_is_my_goal}
\item You are 
%willing to 
about to attempt to 
engage in some joint action\footnote{
We leave open the issue of how joint action is to be characterised subject only to the 
requirement that all joint actions must involve distributive goals (see further footnote \vref{fn:joint_action_distributive_goal}).
Attempts to characterise joint action in ways relevant to explaining development include
	\citet{Tollefsen:2005vh},
	\citet{Carpenter:2009wq},
	\citet{pacherie_framing_2011} and
	\citet{Butterfill:2011fk}.
The last of these 
not only shows how a primitive form of joint action can be characterised without ascriptions of higher-order mental states
but also shares our focus on relations between actions and goals rather than between agents and intentions.
}
or other with me.
%(for example, because you have made eye contact with me while I was in the middle of attempting to do something).

@@

Is it ever possible to know the premises without first knowing the conclusion?  
Consider the first premise.  
Sometimes 
in the right contexts
an individual can recognize in another's facial expressions, engaging gestures or synchronized bodily movements that she is about to attempt to engage in joint action with her.
Exploiting these indicators does not typically depend on knowing the particular contents of any beliefs, desires or goals.  
Expressions, gestures and movements can naturally indicate imminent jointness in much the way they can also naturally indicate emotions.\footnote{
Ideas along these lines are suggested by the discussion of {emergent coordination} in \citet{Knoblich:2010fk}.
}  
Of course these indicators do not guarantee that others are  about to attempt to engage in joint action.  
But they are sufficiently reliable to ground knowledge in some cases.  
The existence of such indicators shows that knowing the first  premise of the above inference does not require already knowing which particular goals the other's actions will be directed to.

Not everything needs to rest on indicators, however.
It is sometimes possible to know that others are about to attempt to engage in joint action with you even without relying on such indicators.  
Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others will act jointly.  
For example, this is often so for children struggling with a coat while surrounded by family or familiar adults.  
And in at least some subcultures people using public transport can reasonably take for granted that, within limits, those around them will act jointly with them when the need is clear.
Of course dispositions to engage in joint action may vary between cultural groups and situations.
This may be fatal for the frequent traveller,
but for others what matters is not whether the dispositions are universal but only that they are sufficiently widespread to be predictable.

Turning to the second point,
could knowing the premises of the your-goal-is-my-goal inference ever put one in a position to know the conclusion?  
Of course the inference is not deductive and will only work when certain background conditions are met.  
These background conditions include the other having largely true beliefs concerning which goal your actions are or will be directed to.
After all, another agent may attempt to act jointly with you while being entirely mistaken about the goals of your actions.
Where this happens, the premises of the inference might be true but the conclusion false.
We shall return to this point later (in section \vref{sec:false_belief}). 
For now, note that in some situations there is no requirement  to consider  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.

As well as having true beliefs concerning the goal of your actions, the other must be willing to pursue this goal. 
This is not guaranteed by the truth of the premises of the inference. 
For all these require, the other might be aware of the goal you envisage but attempting to initiate a joint action directed to another, quite different goal%
%and would be unwilling to perform actions directed to the goal you envisage, so that the conclusion of the inference is false
.  
Perhaps, for example, the other disapproves of your project and is offering to act with you as part of an attempt to divert you.\footnote{
We are grateful to anonymous referee for emphasising this point.
}
Note, however, that it is only reasonable for the other to attempt a joint action with you where you are in a position to know which goal the joint action would be directed to. 
If no alternative goal ought to be salient to you and the other has done nothing to make an alternative salient, then it is not unreasonable for you to ignore the possibility that the other is unwilling to pursue the goal to which your actions will be directed. 
This might easily be the case where, for instance, you are in a tidy kitchen patiently chopping your way through a large pile of carrots; it is perhaps less likely to be the case in a more cluttered kitchen where you are surrounded by all kinds of potential ingredient in different states of preparation. 
In general, how strongly the truth of the premises of the your-goal-is-my-goal inference supports the truth of the conclusion will depend (among other conditions) on how salient alternative goals to which your actions might be directed ought to be to you.

Given these and perhaps other background conditions,
that others are about to attempt to engage in joint action with you 
is sometimes sufficient reason to hold that
you will end up acting jointly with them
 even where the goal to which your actions will be directed is already fixed.

@@

We should also acknowledge that 
mindreaders who only observe can make use of the premise about joint action that your-goal-is-my-goal depends on.  
Knowing that the target of their mindreading  is about to attempt to engage in joint action with another agent might enable them to identify the goal of their target's actions. 
We return to this issue in section \vref{sec:observing_joint_action}. 
For now what matters is just that only interacting mindreaders can exploit your-goal-is-my-goal.

@@

The problem of opaque means was this:
 failures to identify to which ends actions are means can impair goal ascription (see section \vref{sec:opaque_means}).
%As we saw, this problem is particularly likely to arise in cases involving novel tools and communicative gestures.
Showing how your-goal-is-my-goal makes it possible to avoid this problem is a way of demonstrating the potential value   of interaction for mindreading.

In our earlier example 
a novice parent is struggling to lift his heavy twin pram onto a bus when a stranger joins in and they lift the pram together (see page \pageref*{twin_pram}).
Suppose the stranger starts tipping the pram in a way that the novice parent fails to recognise as a means, indeed the only means, of getting it onto the bus.
Outside the context of joint action, 
failure to realise that the action is a means to getting the pram onto the bus
might have the consequence that
the parent's evidence on balance supports the conclusion that the goal of the stranger's actions is to take the pram off the bus.
(Perhaps the stranger is impatient to get onto the bus herself.)
%Or suppose the stranger starts by pulling the pram away from the bus in order to better position it for entry,
%and the novice parent does not realise that this is necessary.
%Outside the context of joint action, 
%his evidence might on balance support the conclusion that the goal of the stranger's actions is to take the pram off the bus.
%(Perhaps the stranger is impatient to get onto the bus herself.)
But in the context of joint action,
your-goal-is-my goal gives the parent  additional evidence for supposing that, 
even though the stranger's actions do not seem to him to be a means to getting the pram onto the bus, 
this really is the goal of her actions.
Of course this additional evidence will not always outweigh   other evidence.
But it will do sometimes, and this is all we need.
We have identified evidence for goal ascription that is available independently of a goal ascriber's knowing which ends actions are means to, and we have illustrated how this will sometimes enable interacting mindreaders to avoid the problem of opaque means
